---
title: "Topic 2: Text Data in R"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Connect to the New York Times API and send a query

```{r load-packages}
library(jsonlite) #convert results from API queries into R-friendly formats 
library(tidyverse) 
library(tidytext) #text data management and analysis
library(ggplot2) #plot word frequencies and publication dates
```

### OK, it's working but we want more data. Let's set some parameters for a bigger query

```{r}
term <- "methane" # Need to use + to string together separate words
begin_date <- "20210120"
end_date <- Sys.Date()
api_key <- 

#construct the query url using API operators
baseurl <- paste0("http://api.nytimes.com/svc/search/v2/articlesearch.json?q=",term,
                  "&begin_date=",begin_date,"&end_date=",end_date,
                  "&facet_filter=true&api-key=",api_key, sep="")

#examine our query url
baseurl
```

```{r, eval=FALSE}
#this code allows for obtaining multiple pages of query results 
 initialQuery <- fromJSON(baseurl)
maxPages <- round((initialQuery$response$meta$hits[1] / 10)-1) 

pages <- list()
for(i in 0:maxPages){
  nytSearch <- fromJSON(paste0(baseurl, "&page=", i), flatten = TRUE) %>% data.frame() 
  message("Retrieving page ", i)
  pages[[i+1]] <- nytSearch 
  Sys.sleep(6) 
}
class(nytSearch)

#need to bind the pages and create a tibble from nytDa


```

```{r article-type}
nytSearch %>% 
  group_by(response.docs.type_of_material) %>%
  summarize(count=n()) %>%
  mutate(percent = (count / sum(count))*100) %>%
  ggplot() +
  geom_bar(aes(y=percent, x=response.docs.type_of_material, fill=response.docs.type_of_material), stat = "identity") + coord_flip()
```

```{r date-plot}
nytSearch %>%
  mutate(pubDay=gsub("T.*","",response.docs.pub_date)) %>%
  group_by(pubDay) %>%
  summarise(count=n()) %>%
  filter(count >= 2) %>%
  ggplot() +
  geom_bar(aes(x=reorder(pubDay, count), y=count), stat="identity") + coord_flip()
```

The New York Times doesn't make full text of the articles available through the API. But we can use the first paragraph of each article.

```{r plot_frequencies}

names(nytSearch)
paragraph <- names(nytSearch)[6] #The 6th column, "response.doc.lead_paragraph", is the one we want here.  
tokenized <- nytSearch %>%
  unnest_tokens(word, paragraph)

tokenized %>%
  count(word, sort = TRUE) %>%
  filter(n > 5) %>% #illegible with all the words displayed
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word)) +
  geom_col() +
  labs(y = NULL)
```

Uh oh, who knows what we need to do here?

```{r}

data(stop_words)

tokenized <- tokenized %>%
  anti_join(stop_words)

tokenized %>%
  count(word, sort = TRUE) %>%
  filter(n > 5) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word)) +
  geom_col() +
  labs(y = NULL)
```

OK, but look at the most common words. Does one stick out?

```{r cleaning, eval=FALSE}

#inspect the list of tokens (words)
tokenized$word

clean_tokens <- str_replace_all(tokenized$word,"land[a-z,A-Z]*","land") #stem tribe words
clean_tokens <- str_remove_all(clean_tokens, "[:digit:]") #remove all numbers
clean_tokens <- str_remove_all(clean_tokens, "washington")
clean_tokens <- gsub("’s", '', clean_tokens)

tokenized$clean <- clean_tokens

tokenized %>%
  count(clean, sort = TRUE) %>%
  filter(n > 10) %>% #illegible with all the words displayed
  mutate(clean = reorder(clean, n)) %>%
  ggplot(aes(n, clean)) +
  geom_col() +
  labs(y = NULL)

#remove the empty strings
tib <-subset(tokenized, clean!="")

#reassign
tokenized <- tib

#try again
tokenized %>%
  count(clean, sort = TRUE) %>%
  filter(n > 10) %>% #illegible with all the words displayed
  mutate(clean = reorder(clean, n)) %>%
  ggplot(aes(n, clean)) +
  geom_col() +
  labs(y = NULL)

```

## Assignment (Due by Week 3)

1.  Create a free New York Times account (<https://developer.nytimes.com/get-started>)

2.  Pick an interesting environmental key word(s) and use the jsonlite package to query the API. Pick something high profile enough and over a large enough time frame that your query yields enough articles for an interesting examination.

3.  Recreate the publications per day and word frequency plots using the first paragraph

-   Make some (at least 3) transformations to the corpus (add stopword(s), stem a key term and its variants, remove numbers)

4.  Recreate the publications per day and word frequency plots using the headlines variable (response.docs.headline.main). Compare the distributions of word frequencies between the first paragraph and headlines. Do you see any difference?